---
title: "MATH 216 Homework 3"
author: "Shannia Fu"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
```



## Admistrative:

Please indicate

* Who you collaborated with: Bianca Gonzalez, Brenda Li
* Roughly how much time you spent on this HW so far: 2.75 hours
* The URL of the RPubs published URL [here](http://rpubs.com/shannia/hw3).
* What gave you the most trouble: not getting enough directions...
* Any comments you have: my knitted file says that the training accuracy is lower than the testing accuracy, but the opposite is true in my r console (despite set seed)




## Question 1:

We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
profiles <- profiles %>% 
  select(-contains("essay"))
```



#### a)

Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.
* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
set.seed(76)

profiles <- profiles %>% 
  mutate(orientation = ifelse(is.na(orientation), "did not report", orientation)) %>% 
  mutate(job = ifelse(is.na(job), "did not report", job))

training_set <- profiles %>% 
  sample_n(0.05*nrow(profiles))
testing_set <- anti_join(profiles, training_set, by="id")

```



#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}


predict_sex_model <- glm(is_female~orientation+job, data=training_set, family="binomial")


training_prediction <- training_set %>% 
  select(id, orientation, job, is_female) %>% 
  mutate(reg_per = fitted(predict_sex_model)) %>% 
  mutate(prediction_female = ifelse(reg_per > 0.5, 1, 0)) %>% 
  mutate(is_corr = ifelse (is_female==prediction_female, 1, 0))

mean(training_prediction$is_corr)

```

The training accuracy is 66.00%.

#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
testing_prediction <- testing_set %>% 
  select(id, orientation, job, is_female) %>% 
  mutate(prediction_female = predict(predict_sex_model, newdata=testing_set, type="response")) %>% 
  mutate(prediction_female = ifelse(prediction_female > 0.5, 1, 0)) %>% 
  mutate(is_corr = ifelse(prediction_female == is_female, 1, 0))

mean(testing_prediction$is_corr)

```

The testing accuracy is 64.65%.

#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

The model performed slightly better on the training data using sexual orientation and job as predictor variables. To be honest neither predictor was a good predictor, as on both the training set and the testing set the "accuracy" of the prediction was around 65%.
The reason for this is probably that because the regression was modeled on the training set, it should perform better on the training set. However, given that the two accuracy levels were around the same, the randomly selected rows used to generate the regression were probably a good representative of the overall data.

(Post-Submission Edit-I ran it with the set.seed(76) line again and the results changed once again--the testing set still tested lower than than the training set, even though it favored the testing set slightly when we did the trials during our discussion.)

## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

bitcoin <- Quandl("BAVERAGE/USD") %>% 
  mutate(prev_day_value = lead(`24h Average`, k = 1)) %>% 
  mutate(volatility = (`24h Average`-prev_day_value)/prev_day_value) %>% 
  filter(!is.na(volatility)) %>% 
  mutate(type="Bitcoin")

gold <- Quandl("BUNDESBANK/BBK01_WT5511") %>%  
  mutate(prev_day_value = lead(Value, k = 1)) %>% 
  mutate(volatility = (Value-prev_day_value)/prev_day_value) %>% 
  filter(!is.na(volatility)) %>% 
  mutate(type="Gold")

bg <- bind_rows(gold, bitcoin) %>% 
  filter(Date %within% interval(ymd("2010-07-18"), ymd("2016-10-31")))

bg_volatility <- ggplot(bg, aes(x=Date, y=volatility*100, col = type)) + geom_line()
bg_volatility +
  labs(title="Volatility of Gold and Bitcoin", x="Date by Year", y="Volatility by Percentage")


```


Bitcoin is much more volatile on a day to day basis compared to gold. My non-econ-educated advice would be to invest in gold to be safe, unless you're very certain about the value growth of bitcoin.


## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}

newjuke <- jukebox %>% 
  mutate(date_time = parse_date_time(date_time, "a m d HMS Y")) %>% 
  filter(hour(date_time) >= 0 & hour(date_time) <= 8) %>% 
  filter(month(date_time) >= 9 | month(date_time) <= 5) %>% 
  group_by(artist) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  head(10)

knitr::kable(newjuke)

```

These are the top ten bands that played at Reed during the graveyard shift throughout the academic year.


